### 2018-06-20 Writing/response_to_reviewers.tex

\item{\textcolor{blue}{"However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
I'm not sure I follow this.}}
\label{dishonest}

We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is not at all a good practice in systematics (it can be seen as dishonest to simply remove characters because they ``don't give the right signal''). We've removed the end of this sentence however to avoid any further confusion.

\item{\textcolor{blue}{"However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree. "\\
I understand why you did this, but I think you should also compare normal to true. That will give us some idea of the scope of how close we can get to true, under these conditions, and will help contextualize the results.}}

We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 9 to 11.
We've added a reference to these figures in the main text:

\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.} lines @@@

@@

\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>$ 0.95 and different if they had an overlap probability $>$ 0.05.''}}

We fixed this type. We considered them different if they had an overlap probability $<$ 0.05.

\item{\textcolor{blue}{It is not clear how the conclusions were reached. For example, line 381-384: ``Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table 5 - median NTSRF of 0:828 and 0:679 respectively) but not in terms of individual taxon placements (Table 5 - median NTSTr of 0:738 and 0:601 respectively).'' Why is the 0:828 - 0:679 = 0:149 significant for this test and 0:738 - 0:601 = 0:137 is not?}}

The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 5.
We now detailed the values and where they are present for each of the tests to avoid confusion.

### 2018-06-20 Writing/draft_highlighted.tex
updated table and figures order

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.}

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 1}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 1}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 1}).
There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 2}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 2}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.

@@

Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 3}).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 4}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 4}).

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).

>>>>

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 4 and 5).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.}

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 2}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 2}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 2}).
There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 3}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.

@@

Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 4}).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 5}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 5}).

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 6, 7 and 8}).
